\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-09 k-planar.tex
continuiing on appendix

By removing both spars from each kite of $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three edges or $4$ edges; see \cref{one_planar_example}(b).  Observe that, for any spar $xy\in E(G)\setminus E(G')$ that crosses $vw\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).  Therefore $\dist_{G'}(v,w)\le 2$.

Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}.
\begin{lem}
	\label{induction} The setup:
	\begin{compactenum}
		\item Let $G$ and $G'$ be defined as above.
		\item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that
		\begin{compactenum}
			\item No edge of $F$ is crossed by any edge of $G$; and
			\item $V(F)$ can be partitioned into $P_1,\ldots,P_k$, for some $k\in\{1,2,3\}$ such that for each $i\in\{1,\ldots,k\}$,
			\begin{compactenum}
				\item $F[P_i]$ is a path; and
				\item $|V(P_i)\cap L_j| \le 3$ for every integer $j\ge 0$.
			\end{compactenum}
		\end{compactenum}
% \note{DW}{Should the vertices already grabbed by $P_1,P_2,P_3$ be mentioned in the setup? The new paths $Q_1,Q_2,Q_3$ need to avoid the vertices already grabbed by $P_1,P_2,P_3$. Basically, I am concerned about what happens when distinct vertical paths in $T$ want to grab the same vertex.} \note{PM}{No, at this point we're only promising an $H$-partition of $N$, the part of $G$ contained in the uncrossed cycle $F$.}
		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $F$ or the interior of $F$.
	\end{compactenum}
	Then $N$ has an $H$-partition $\PP=\{S_x\cup\{v_x\} : x\in V(H)\}$ such that:
	\begin{compactenum}
		\item $H$ is planar;
		\item for every integer $j\ge 0$ and every $x\in V(H)$, $|S_x\cap L_j|\le 3$;
		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; 
%		\note{DW}{Should this be $P_i\subseteq S_{x_i}$? (since the path grabs other vertices into its part)}\note{PM}{No, for the same reason as above.}
 and
		\item $H$ has a tree-decomposition in which every bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
	\end{compactenum}
\end{lem}

@@

  If $k\in\{1,2\}$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge of $F$ on its boundary.  If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.  (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.  Now add edges to the inner faces of $N'$ to obtain a supergraph $N''$ in which each inner face has three $3$ edges on its boundary.
  By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.  Let $\tau:=v_1,\ldots,v_p$ be the face of $N'$ that contains $\tau''$.  (Note that $p\in\{3,4\}$.)

@@

  For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to $V(F)$.
   %
   %
   % If $k=3$, set $R_i := P_i$ for each $i\in\{1,2,3\}$.  Otherwise, as in \citep{DJMMUW20}, split $P_1,\ldots,P_k$ to partition $V(F)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $F[R_i]$ is a non-empty path and each $R_i$ contains vertices from exactly one of $P_1,\ldots,P_k$.
  %
	% Next, as in \citep{DJMMUW20}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $N'-V(F)$.
  Let $\overline{Y}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.  Let $S:=V(Q_1)\cup V(Q_2)\cup V(Q_3)$ and let $v:=v_p$.  Observe that, for each $i\in\{1,2,3\}$ and each $j\ge 0$, $|V(Q_i)\cap L_j|\le 1$.  Therefore $|S\cap L_j|\le 3$ for each integer $j\ge 0$.  Therefore, taking $Y:=S\cup\{v\}$ as a part in the partition $\PP$ will satisfy Requirement~2 of the lemma.

	% We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap L_i|\le 15$.  First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most three vertices of $L_i$, each incident on at most two edges of $\overline{Y}$.  Since $\dist_{G'}(v,w)\le 2$ for each $vw\in E(G)$, any vertex $x\in V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i$, is incident to an edge $xy\in E(G)$ that crosses one of the at most six edges in $\overline{Y}$ having an endpoint in $L_i$.  These at most six edges have at most 12 endpoints.  Therefore $|V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i|\le 6\times 2=12$, so $|V(\overline{Y}^+)\cap L_i|\le 12+3=15$.

	Let $M$ denote the subgraph of $G'$ containing the edges and vertices of $\overline{Y}$ and the edges and vertices of $F$. (See \cref{boring_figure}.)  By our choice of $\tau$, the graph $M$ is $2$-connected.  Furthermore, the vertices on the boundary of each inner face of $M$ are contained in $Y\cup P_a\cup P_b$ for some $a,b\in\{1,\ldots,k\}$.  In particular if $k=3$ then, for each inner face $F_i$ of $M$ there exists some $c\in\{1,2,3\}$ such that no vertex of $P_c$ is on the boundary of $F_i$.

  \begin{figure}
    \begin{center}
      \begin{tabular}{c@{\hspace{1cm}}c}
        \includegraphics{figs/mmplus-1} & \includegraphics{figs/mmplus-2} \\
        $M$ & $M^+$
      \end{tabular}
    \end{center}
    \caption{The graphs $M$ and $M^+$. \note{PM}{This figure needs updating now.}}
    \label{boring_figure}
  \end{figure}

  The graph $M$ has some number of inner faces $F_1,\ldots,F_m$, each of which is contained in the interior of $F$. 
  % Some of the inner faces of $M^+$ are Type~(ii) faces of $G$. Let $F_1,\ldots,F_m$ be the inner faces of $M^+$ that are not Type~(ii) faces of $G$.  We claim that, for each $i\in\{1,\ldots,m\}$, the boundary of $F_i$ is a cycle in $G'$ that contains no spars. Otherwise, some edge $vw$ contributes to the boundary of $F_i$ but is crossed by an edge $xy\in E(G)$. Then, $vw\not\in E(F)$ since no edge of $F$ is crossed by any edge of $G$. Therefore $vw\in E(\overline{Y}^+)$ so $xy\in E(\overline{Y}^+)$. But then the only faces of $M^+$ with $vw$ on their boundary are Type~(ii) faces.  In particular $vw$ cannot be on the boundary of $F_i$.
  Each of the faces $F_1,\ldots,F_m$ is contained in a single inner face of $M$.  Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.  Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(F_i)$ to obtain the desired $H$-partition $\PP_i$ and tree decomposition of subgraph $N_i$ of $G$ contained in $F_i$.

@@

\begin{lem}
Fix an integer $d \geq 3$. Say $G_0$ is a 2-connected plane multigraph, in which each facial cycle has length in $\{3,\dots,d\}$. \note{DW}{I expect $d$ will be irrlevant to this lemma.} Let $T$ be a spanning tree of $G_0$ rooted at some vertex $r$ on the boundary of the outerface of $G_0$. Let $k\in\{1,2,3\}$ and let $C = [A_1,\dots,A_k]$ be a cycle in $G_0$ with $r$ is in the exterior of $C$, where 
$A_1, \dots, A_k$ are bipods in $G_0$. Let $G$ be the plane graph consisting of all the edges and vertices of $G_0$ in $C$ and the interior of $C$.
Then $G$ has a partition $\PP$ into tripods with resect to $T$ such that $P_1,\dots,P_k\in\PP$ and the graph $H : = G / \PP$ is planar and has a tree-decomposition in which every bag has size at most 4 and some bag contains all the vertices of $H$ corresponding to $P_1,\dots,P_k$. \note{DW}{Do we need to add that ``each face of $G_0$ is the centre of some tripod in $\PP$'' or something like that? Then each facial clique will be in some part.}
\end{lem}

\begin{proof}
	\note{DW}{Is $C$ allowed to be a 2-cycle?}
The proof is by induction on $|V(G)\setminus V(C)|$. If $V(G)\setminus V(C)=\emptyset$, then the desired partition is $\PP=\{P_1,\ldots,P_k\}$.  The tree-decomposition of $H$ consists of a single bag that contains the $k\le 3$ vertices corresponding to $P_1,\ldots,P_k$.

@@

Let $G'$ be an arbitrary triangulation of $G_0$. For $i\in\{1,2,3\}$, colour each vertex in $R_i$ by $i$. Now, for each remaining vertex $v$ in $G$, consider the path $P_v$ from $v$ to the root of $T$. Since $r$ is on the outer-face of $G_0$, $P_v$ contains at least one vertex of $C$. If the first vertex of $P_v$ that belongs to $F$ is in $R_i$ then assign the colour $i$ to $v$. We obtain a 3-colouring of the vertices of $G'$ that satisfies the conditions of Sperner's Lemma. By Sperner's Lemma there exists a triangular face $\tau=v_1v_2v_3$ of $G'$ whose vertices are coloured $1,2,3$ respectively. Let $F$ be the face of $G_0$ containing $\tau$.

For each $i\in\{1,2,3\}$, let $Q_i$ be the path in $T$ from $v_i$ to the first ancestor $v_i'$ of $v_i$ in $T$ that is contained in $C$. Observe that $Q_1$, $Q_2$, and $Q_3$ are disjoint since $Q_i$ consists only of vertices coloured $i$. Note that $Q_i$ may consist of the single vertex $v_i=v_i'$.  Let $Q_i'$ be $Q_i$ minus its final vertex $v_i'$.  Imagine for a moment that the cycle $C$ is oriented clockwise, which defines an orientation of $R_1$, $R_2$ and $R_3$. Let $R_i^-$ be the subpath of $R_i$ that contains $v'_i$ and all vertices that precede it, and let $R_i^+$ be the subpath of $R_i$ that contains $v'_i$ and all vertices that succeed it.

@@

, where we use the convention that $Q_4=Q_1$ and $R_4=R_1$. Note that $C_i$ may be \emph{degenerate} in the sense that $[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$ may consist only of a single edge $v_iv_{i+1}$.

Let $Y$ be the tripod consisting of $Q_1'\cup Q_2' \cup Q_3'$ plus the vertices and edges of $F$ between non-empty $Q_1',Q_2',Q_3'$.

\note{DW}{Much of the following paragraph can be deleted} Consider any non-degenerate $C_i=[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$. Note that these four paths  are pairwise disjoint, and thus $C_i$ is a cycle. If  $Q_i'$ and $Q_{i+1}'$ are non-empty, then each is a vertical path in $T$.  Furthermore, each of $R_i^-$ and $R_{i+1}^+$ consists of at most two vertical paths in $T$.  Thus, $C_i$ is the concatenation of at most six vertical paths in $T$ \note{DW}{re-do}.  Let $G_i$ be the subgraph of $G_0$ consisting of all the edges and vertices of $G_0$ contained in $C_i$ and the interior of $C_i$.  Observe that $G_i$ contains $v_i$ and $v_{i+1}$ but not the third vertex of $\tau$.

Consider the graph consisting of the edges and vertices of $F$, the edges and vertices of $C$, and the edges and vertices of $Q_1,Q_2,Q_3$. This graph has up to three \note{DW}{can be more} internal faces $C_1,C_2,C_3$ where each $C_i=[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$ and $R_{i}^+$ and $R_{i}^-$ are the same portions of $R_i$ as defined in \cref{NearTriang} \note{DW}{move in here}    Observe that  $C_i=[ R_i^+ , R_{i+1}^-, I_i ]$, where $R_i^+$ and $R_{i+1}^-$ are bipods, and $I_i$ is the bipod $Q_i' S_i Q_{i+1}'$. Let $G_i$ be the subgraph of $G$ whose vertices and edges are in $C_i$ or its interior.

For $i\in\{1,2,3\}$, if $C_i$ is non-empty, then $G_i$ and $C_i=[ R_i^+ , R_{i+1}^-, I_i ]$ satisfy the conditions of the lemma, and we may apply induction to $G_i$.  (Note that one or two of $R_i^+$, $R_{i+1}^-$ and $I_i$ may be empty, in which case we apply the inductive hypothesis with $k=2$ or $k=1$, respectively.)\  This gives a partition $\PP_i$ of $G_i$ such that $H_i:= G_i / \PP_i$ satisfies the conclusions of the lemma. Let $(B^i_x:x\in V(J_i))$ be a tree-decomposition of $H_i$, in which every bag has size at most 4, and some bag $B^i_{u_i}$ contains the vertices of $H_i$ corresponding to  $R_i^+$, $R_{i+1}^-$ and $I_i$ (if they are non-empty).  We do this for each non-degenerate $C_i$.

>>>>

By removing one spar from each kite of $G$ we obtain a plane multigraph $G'$ each of whose faces is bounded by three edges; see \cref{one_planar_example}(b).  Observe that, for any spar $xy\in E(G)\setminus E(G')$ that crosses $vw\in E(G')$, $G'$ contains the path $vxw$ (and $vyw$).  Therefore $\dist_{G'}(v,w)\le 2$.

Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}.
\begin{lem}
	\label{induction} The setup:
	\begin{compactenum}
		\item Let $G$ and $G'$ be defined as above.
		\item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that
		\begin{compactenum}
			\item No edge of $F$ is crossed by any edge of $G$; and
			\item $V(F)$ can be partitioned into $P_1,\ldots,P_k$, for some $k\in\{1,2,3\}$ such that for each $i\in\{1,\ldots,k\}$,
			\begin{compactenum}
				\item $F[P_i]$ is a path; and
				\item $|V(P_i)\cap L_j| \le 15$ for every integer $j\ge 0$.
			\end{compactenum}
		\end{compactenum}
% \note{DW}{Should the vertices already grabbed by $P_1,P_2,P_3$ be mentioned in the setup? The new paths $Q_1,Q_2,Q_3$ need to avoid the vertices already grabbed by $P_1,P_2,P_3$. Basically, I am concerned about what happens when distinct vertical paths in $T$ want to grab the same vertex.} \note{PM}{No, at this point we're only promising an $H$-partition of $N$, the part of $G$ contained in the uncrossed cycle $F$.}
		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $F$ or the interior of $F$.
	\end{compactenum}
	Then $N$ has an $H$-partition $\PP=\{S_x : x\in V(H)\}$ such that:
	\begin{compactenum}
		\item $H$ is planar;
		\item for every integer $j\ge 0$ and every $x\in V(H)$, $|S_x\cap L_j|\le 15$;
		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; 
%		\note{DW}{Should this be $P_i\subseteq S_{x_i}$? (since the path grabs other vertices into its part)}\note{PM}{No, for the same reason as above.}
 and
		\item $H$ has a tree-decomposition in which every bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
	\end{compactenum}
\end{lem}

@@

  If $k\in\{1,2\}$ then we let $\tau:=v_1v_2v_3$ be an inner face of $N$ that contains at least one edge of $F$ on its boundary.  If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.  (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.   By Sperner's Lemma there exists an inner face $\tau:=v_1v_2v_3$ of $N$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.

@@

  For each $i\in\{1,2,3\}$, let $Q_i$ be the shortest path, in $T$, from $v_i$ to $V(F)$.
   %
   %
   % If $k=3$, set $R_i := P_i$ for each $i\in\{1,2,3\}$.  Otherwise, as in \citep{DJMMUW20}, split $P_1,\ldots,P_k$ to partition $V(F)$ into three sets $R_1$, $R_2$, and $R_3$ such that each $F[R_i]$ is a non-empty path and each $R_i$ contains vertices from exactly one of $P_1,\ldots,P_k$.
  %
	% Next, as in \citep{DJMMUW20}, use Sperner's Lemma to find an inner face $\tau=v_1v_2v_3$ of $N'$ such that, $T$ contains disjoint vertical paths $Q_1,Q_2,Q_3$ such that each $Q_i$ begins at $v_i$, ends at some vertex in $R_i$, and whose internal vertices (if any) are contained in $N'-V(F)$.
	Let $\overline{Y}$ denote the subgraph of $N'$ consisting of vertices and edges of $Q_1$, $Q_2$, $Q_3$, and $\tau$.  Let $\overline{Y}^+$ denote the subgraph of $N$ consisting of the vertices and edges of $\overline{Y}$ plus the vertices and edges of every kite formed by a crossing between an edge of $G$ and an edge of $\overline{Y}$.  Let $Y:= V(\overline{Y}-F)$ and let $Y^+:=V(\overline{Y}^+-F)$.

	We claim that, for each integer $i\ge 0$, $|V(\overline{Y}^+)\cap L_i|\le 15$.  First observe that, since $Q_1,Q_2,Q_3$ are each vertical paths in $T$,  $\overline{Y}$ contains at most three vertices of $L_i$, each incident on at most two edges of $\overline{Y}$.  Since $\dist_{G'}(v,w)\le 2$ for each $vw\in E(G)$, any vertex $x\in V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i$, is incident to an edge $xy\in E(G)$ that crosses one of the at most six edges in $\overline{Y}$ having an endpoint in $L_i$.  These at most six edges have at most 12 endpoints.  Therefore $|V(\overline{Y}^+)\setminus V(\overline{Y})\cap L_i|\le 6\times 2=12$, so $|V(\overline{Y}^+)\cap L_i|\le 12+3=15$.

	Let $M$ and $M^+$ denote the subgraph of $G$ containing the edges and vertices of $\overline{Y}$, respectively $\overline{Y}^+$, and the edges and vertices of $F$. (See \cref{boring_figure}.)  By our choice of $\tau$, the graph $M$ is $2$-connected.  Furthermore, the vertices on the boundary of each inner face of $M$ are contained in $Y\cup P_a\cup P_b$ for some $a,b\in\{1,\ldots,k\}$.  In particular if $k=3$ then, for each inner face $F_i$ of $M$ there exists some $c\in\{1,2,3\}$ such that no vertex of $P_c$ is on the boundary of $F_i$.

  \begin{figure}
    \begin{center}
      \begin{tabular}{c@{\hspace{1cm}}c}
        \includegraphics{figs/mmplus-1} & \includegraphics{figs/mmplus-2} \\
        $M$ & $M^+$
      \end{tabular}
    \end{center}
    \caption{The graphs $M$ and $M^+$.}
    \label{boring_figure}
  \end{figure}

  The graph $M^+$ has some number of inner faces, all contained in the interior of $F$. Some of the inner faces of $M^+$ are Type~(ii) faces of $G$. Let $F_1,\ldots,F_m$ be the inner faces of $M^+$ that are not Type~(ii) faces of $G$.  We claim that, for each $i\in\{1,\ldots,m\}$, the boundary of $F_i$ is a cycle in $G'$ that contains no spars. Otherwise, some edge $vw$ contributes to the boundary of $F_i$ but is crossed by an edge $xy\in E(G)$. Then, $vw\not\in E(F)$ since no edge of $F$ is crossed by any edge of $G$. Therefore $vw\in E(\overline{Y}^+)$ so $xy\in E(\overline{Y}^+)$. But then the only faces of $M^+$ with $vw$ on their boundary are Type~(ii) faces.  In particular $vw$ cannot be on the boundary of $F_i$.

  Observe that each of the faces $F_1,\ldots,F_m$ is contained in a single inner face of $M$.  Therefore, $V(F_i)$ can be partitioned into at most three sets $P_1'$, $P_2'$, and $P_3'$ where $P_1'\subset V(Y^+)$, $P_2'\subseteq P_a$, $P_3'\subseteq P_b$ for some $a,b\in\{1,2,3\}$, and $F_i[P_j']$ is a path, for each $j\in\{1,2,3\}$.  Therefore, we can apply induction using the cycle $F_i$ and the partition $P_1',P_2',P_3'$ of $V(F_i)$ to obtain the desired $H$-partition $\PP_i$ and tree decomposition of subgraph $N_i$ of $G$ contained in $F_i$.

@@

\begin{lem}
Say $G_0$ is a 2-connected plane multigraph. Let $T$ be a spanning tree of $G_0$ rooted at some vertex $r$ on the boundary of the outerface of $G_0$. Let $k\in\{1,2,3\}$ and let $C = [A_1,\dots,A_k]$ be a cycle in $G_0$ with $r$ in the exterior of $C$, where $A_1, \dots, A_k$ are bipods in $G_0$. Let $G$ be the plane graph consisting of all the edges and vertices of $G_0$ in $C$ and the interior of $C$. Then $G$ has a partition $\PP$ into tripods with respect to $T$ such that:
\begin{compactitem}
\item $P_1,\dots,P_k\in\PP$,
\item the quotient $G / \PP$ is planar and has a tree-decomposition $\mathcal{T}$ of width at most 3 such that:
\begin{compactitem}
	\item some bag of $\mathcal{T}$ contains $P_1,\dots,P_k$, and 
	\item for every face $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
\end{compactitem}
\end{compactitem}
\end{lem}

\begin{proof}
	\note{DW}{Can we always assume $k=3$ by allowing $A_i=\emptyset$? That would clean it up.}

	\note{DW}{Is $C$ allowed to be a 2-cycle?}
The proof is by induction on $|V(G)\setminus V(C)|$. If $V(G)\setminus V(C)=\emptyset$, then $G$ has just one face bounded by $C$ and the desired partition is $\PP=\{P_1,\ldots,P_k\}$.  The tree-decomposition of $G/\PP$ consists of a single bag that contains the $k\le 3$ vertices corresponding to $P_1,\ldots,P_k$.

@@

Let $G'$ be a near-triangulation obtained from $G$ by triangulating every internal face. For $i\in\{1,2,3\}$, colour each vertex in $R_i$ by $i$. Now, for each remaining vertex $v$ in $G$, consider the path $P_v$ from $v$ to the root of $T$. Since $r$ is on the outer-face of $G_0$, $P_v$ contains at least one vertex of $C$. If the first vertex of $P_v$ that belongs to $F$ is in $R_i$ then assign the colour $i$ to $v$. We obtain a 3-colouring of the vertices of $G'$ that satisfies the conditions of Sperner's Lemma. By Sperner's Lemma there exists a triangular face $\tau=v_1v_2v_3$ of $G'$ whose vertices are coloured $1,2,3$ respectively. Let $F$ be the face of $G_0$ containing $\tau$.

For each $i\in\{1,2,3\}$, let $Q_i$ be the path in $T$ from $v_i$ to the first ancestor $v_i'$ of $v_i$ in $T$ that is contained in $C$. Observe that $Q_1$, $Q_2$, and $Q_3$ are disjoint since $Q_i$ consists only of vertices coloured $i$. Let $Q_i'$ be $Q_i$ minus its final vertex $v_i'$. Note that $Q_i$ may consist of the single vertex $v_i=v_i'$, in which case $Q_i'$ is empty. Imagine for a moment that the cycle $C$ is oriented clockwise, which defines an orientation of $R_1$, $R_2$ and $R_3$. Let $R_i^-$ be the subpath of $R_i$ that contains $v'_i$ and all vertices that precede it, and let $R_i^+$ be the subpath of $R_i$ that contains $v'_i$ and all vertices that succeed it.

Let $Y$ be the tripod consisting of $Q_1'\cup Q_2' \cup Q_3'$ plus the vertices and edges of $F$.

@@

Let $L_i$ be the path $Q_i'S_iQ_{i+1}'$ for $i\in\{1,\dots,k\}$. So $L_i$ is a bipod.

, where we use the convention that $Q_4=Q_1$ and $R_4=R_1$.

%Note that $C_i$ may be \defin{degenerate} in the sense that $[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$ may consist only of a single edge $v_iv_{i+1}$.

% Consider any non-degenerate $C_i=[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$. Note that these four paths  are pairwise disjoint, and thus $C_i$ is a cycle. If  $Q_i'$ and $Q_{i+1}'$ are non-empty, then each is a vertical path in $T$.  Furthermore, each of $R_i^-$ and $R_{i+1}^+$ consists of at most two vertical paths in $T$.  Thus, $C_i$ is the concatenation of at most six vertical paths in $T$ \note{DW}{re-do}.  Let $G_i$ be the subgraph of $G_0$ consisting of all the edges and vertices of $G_0$ contained in $C_i$ and the interior of $C_i$.  Observe that $G_i$ contains $v_i$ and $v_{i+1}$ but not the third vertex of $\tau$.

Consider the plane graph $Z$ consisting of the edges and vertices of $F$, the edges and vertices of $C$, and the edges and vertices of $Q_1,Q_2,Q_3$. Then $Z$ is 2-connected, each face of $Z$ is bounded by a cycle, and the outerface of $Z$ is $C$. Let $Z_1,\dots,Z_m$ be the internal faces of $Z$.

Consider $j\in\{1,\dots,m\}$. Let $G_j$ be the subgraph of $G$ consisting of all the edges and vertices of $G$ in $Z_j$ and the interior of $Z_j$. The outerface of $G_j$ is $Z_j$. Observe that for some $i\in\{1,\dots,k\}$, we have
$Z_j=[R_{i,j}^+,R_{i+1,j}^-,S_{i,j}]$, where 
$R_{i,j}^+$ is some subpath of $R_i^+$,
$R_{i+1,j}^-$ is some subpath of $R_{i+1}^-$, and
$S_{i,j}$ is some subpath of $S_i$.
Since $R_i^+$, $R_{i+1}^-$ and $S_i$ are bipods, 
$R_{i,j}^+$, $R_{i+1,j}^-$ and $S_{i,j}$ are bipods. 
By construction, $v_{i-1}$ is not in $G_j$. 
Thus $V(G_j)\setminus V(Z_j) \subsetneq V(G)\setminus V(C)$, and we may apply induction to $G_j$.  Thus $G_j$ has a partition $\PP_j$ into tripods with respect to $T$ such that:
\begin{compactitem}
	\item $R_{i,j}^+, R_{i+1,j}^-, S_{i,j}\in\PP_j$,
	\item the quotient $G_j / \PP_j$ is planar and has a tree-decomposition $\mathcal{T}_j$ of width at most 3 such that: 
	\begin{compactitem}
		\item some bag $B_j$ of $\mathcal{T}_j$ contains $R_{i,j}^+, R_{i+1,j}^-, S_{i,j}$, and 
		\item for every face $F$ of $G_j$ there is a bag $B$ of $\mathcal{T}_j$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
	\end{compactitem}
\end{compactitem}
\end{lem}

%This graph has up to three \note{DW}{can be more} internal faces $C_1,C_2,C_3$ where each $C_i=[Q_i',R_i^+,R_{i+1}^-,Q_{i+1}']$ and $R_{i}^+$ and $R_{i}^-$ are the same portions of $R_i$ as defined in \cref{NearTriang} \note{DW}{move in here}    Observe that  $C_i=[ R_i^+ , R_{i+1}^-, I_i ]$, where $R_i^+$ and $R_{i+1}^-$ are bipods, and $I_i$ is the bipod $Q_i' S_i Q_{i+1}'$. Let $G_i$ be the subgraph of $G$ whose vertices and edges are in $C_i$ or its interior.

%For $i\in\{1,2,3\}$, if $C_i$ is non-empty, then $G_i$ and $C_i=[ R_i^+ , R_{i+1}^-, I_i ]$ satisfy the conditions of the lemma, and we may apply induction to $G_i$.  (Note that one or two of $R_i^+$, $R_{i+1}^-$ and $I_i$ may be empty, in which case we apply the inductive hypothesis with $k=2$ or $k=1$, respectively.)\  This gives a partition $\PP_i$ of $G_i$ such that $H_i:= G_i / \PP_i$ satisfies the conclusions of the lemma. Let $(B^i_x:x\in V(J_i))$ be a tree-decomposition of $H_i$, in which every bag has size at most 4, and some bag $B^i_{u_i}$ contains the vertices of $H_i$ corresponding to  $R_i^+$, $R_{i+1}^-$ and $I_i$ (if they are non-empty).  We do this for each non-degenerate $C_i$.
